\chapter{Evaluation}
\label{sec:evaluation}

%N-fold cross-validation with flat prior, complexity, depth, pcfg or temperley's prior
%Use different likelihoods: Additive noise, downbeat stretching, louder downbeats

This chapter will first give an overview of how we will evaluate the parser. The rest of this chapter is structured as follows: An overview of the evaluation criteria is given in section \ref{sec:criteria} and the evaluation measure will be explained in section \ref{sec:measure}. 

The parser produces a ranked list of hypotheses. The rhythmic analysis $R$ that corresponds to the most highly ranked hypothesis is assumed to be the parser's interpretation of its input. A parser output $R$ will be evaluated against a gold-standard parse $R^*$ from the corpus by an evaluation measure function. 

We cannot afford to keep a completely separate test set since our corpus is fairly small. Therefore we will train the models on different training sets and evaluate them on small parts of the corpus left out of the training set. This process is known as cross-validation. For $n$-fold cross-validation, the corpus is divided into $n$ equal parts, a training set is constructed from $n-1$ of these parts and a test set is constructed of one of these parts. This is done $n$ times and the parts sampled randomly from the corpus. When dividing into $n$ random sampled parts, performances are treated as whole units. Since there are twenty different performances in the corpus, using 10-fold cross validation results in training sets of eighteen performances that will be evaluated on test sets of two performances.

Furthermore, some performances contain over a hundred onsets. The parser is too slow to parse performances longer than around thirty onsets. Therefore the parser will be tested on small slices of performances, consisting of the first few notes of a performance with some preferred length.

Our subdivision trees can only represent performances with a length, measured in whatever metrical unit, that is a power of two. If a performance is for example three bars long, the subdivision tree will represent a fourth bar as a tie. To make evaluation simpler, the test performances are restricted to have lengths that are a power of two. This is done by selecting, given some preferred length, the leftmost nested subdivision tree that contains a number of onsets closest to the preferred length. To do this, we use the subdivision trees in our corpus.

\section{Criteria}
\label{sec:criteria}

To assess the quality of a parser output we will look at how many properties of the gold-standard rhythm were analysed correctly and how many properties of the parse occur in the gold standard rhythm. The subdivision trees describe two properties of rhythm: the location of the down- and upbeats, which is also called the \textit{phase}, and the subdivision pattern, which provides information about the time signature. The evaluation measure should somehow measure the extend to which the parser output is consistent with the gold standard in terms of subdivision and down- and upbeat locations. 

A parser output may be consistent with some aspects of the gold-standard but inconsistent with others. In figure \ref{fig:div_error} for example, the parser correctly identified the downbeat but incorrectly assumed a triple division. It is also possible that the parser correctly identifies the divisions but incorrectly identifies the phase, examples of these kind of errors are shown in figure \ref{fig:phase_error}. Note that a phase error at the deepest level is more severe than a phase error at a higher level. If the phase is incorrect at the deepest level, every downbeat will be identified as an upbeat and vice-versa. If the phase is incorrect at a higher level, the down- and upbeats at the lowest level are still correct. This can be understood intuitively as well: a phase error at the lowest level is more severe since it makes the entire rhythm syncopated, for a phase error at a higher level, the notes on the lower levels would still be in phase.

Notably, figure \ref{fig:phase_error} introduces rests in the parse trees. Rests are necessary to represent that the last note is a quarter note and not a whole note. The parser introduced in in this thesis \ref{sec:method} does not handle rests. This issue is discussed further in chapter \ref{sec:rests}.
\begin{figure}
\centering
\subfloat[Gold-standard.]{
\parbox[3cm]{0.5\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .$\bullet$ ] [ .$\bullet$ ] ]
}
}
\subfloat[Parser output.]{
\parbox[3cm]{0.5\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .$\bullet$ ] [ .$*$ ] [ .$\bullet$ ] ] 
}
}
\caption{An example of incorrect division detection.}
\label{fig:div_error}
\end{figure}


\begin{figure}
\subfloat[Gold-standard.]{
\label{fig:precision_error:a}
\parbox{0.33\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] ]
}
}
\centering
\subfloat[]{
\parbox{0.33\linewidth}{
\label{fig:precision_error:b}
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] [ .{$\frac{1}{2}$} [ .{$\frac{1}{4}$} [ .$*$ ] [ .$\bullet$ ] ] [ .$\bullet$ ] ] ]
}
}
\subfloat[]{
\label{fig:precision_error:c}
\parbox{0.33\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .{$\frac{1}{4}$} [ .{$\frac{1}{8}$} [ .$*$ ] [ .$\bullet$ ] ] [ .$*$ ] ] ] [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] ]
}
}
\caption{An example of too detailed analyses (resulting in a lower precision)}
\label{fig:precision_error}
\end{figure}

\begin{figure}
\subfloat[]{
\parbox{0.33\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] ]
}
}
\subfloat[Gold-standard.]{
\parbox{0.33\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] [ .{$\frac{1}{2}$} [ .{$\frac{1}{4}$} [ .$*$ ] [ .$\bullet$ ] ] [ .$\bullet$ ] ] ]
}
}
\centering
\caption{An example of too simple analyses (resulting in a lower recall).}
\label{fig:recall_error}
\end{figure}



\begin{figure}[t]
\subfloat[Gold-standard.]{
\label{fig:phase_error:a}
\parbox{0.2\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] ]
}
}
\centering
\subfloat[Phase error at the half note level]{
\label{fig:phase_error:b}
\parbox{0.3\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$*$ ] [ .{$\frac{1}{4}$} [ .$\bullet$ ] [ .$\bullet$ ] ] ] [ .{$\frac{1}{2}$} [ .{$\frac{1}{4}$} [ .$\bullet$ ] [ .$\bullet$ ] ] [ .$*$ ] ] ]
}
}
\subfloat[Lowest level phase error (most severe).]{
\label{fig:phase_error:c}
\parbox{0.5\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .{$\frac{1}{4}$} [ .$*$ ] [ .$\bullet$ ] ] [ .{$\frac{1}{4}$} [ .$\bullet$ ] [ .$\bullet$ ] ] ] [ .{$\frac{1}{2}$} [ .{$\frac{1}{4}$} [ .$\bullet$ ] [ .Rest ] ] [ .Rest ] ] ]
}
}
\caption{An example incorrect phase detection. See section \ref{sec:discussion} for why there are rests in this subdivision tree.}
\label{fig:phase_error}
\end{figure}

\section{Evaluation Measure}
\label{sec:measure}

The precision and recall of the parser are measured as follows: The precision is the number of subdivisions and down- and upbeats in the parser output that were correctly identified with respect to the gold-standard, divided by the total number of subdivisions and down- and upbeats in the parser output. The recall is the number of subdivisions and down- and upbeats in the gold-standard that are correctly identified by the parser output, divided by the total amount of subdivisions and down- and upbeats in the gold-standard.

To measure these quantities, every onset governed by $R$ and $R^*$ is converted to a list of claims that this onset makes about the structure. For example, the first quarter note of a 4/4 measure claims to be a downbeat at the half note level and a downbeat at the quarter note level (see for example figure \ref{fig:phase_error:a}). The second quarter note claims to be governed by a downbeat at the half note level and to be an upbeat at the quarter note level. The third onset claims to be an upbeat at the half note level and a downbeat at the quarter note level. 

An evaluation that uses these kinds of lists penalises phase errors at the lower levels more severe than phase errors at higher levels. Given the gold-standard in figure \ref{fig:phase_error:a}, consider the following hypothetical errors in the parser output: Should the phase be wrong at the quarter note level (figure \ref{fig:phase_error:b}), a downbeat at the quarter note level becomes  an upbeat, however, the downbeat would still be governed by the downbeat at the half note level. Should the phase be wrong at the half note level (figure, \ref{fig:phase_error:c}), the down- and upbeats at the quarter note level are still correct.

There are two issues with the evaluation as suggested above. First, since every onset lists all claims it implies at higher levels, the parser will get points for divisions at higher levels multiple times. Second, figure \ref{fig:phase_error} shows that a phase error can lead to another level to be added to the tree, which introduces an inconsistency in the interpretation of levels between the parse tree and the gold-standard. 

The first issue is remedied by keeping a list of parser decisions that have been accounted for. If the first onset in figure \ref{fig:phase_error:a} claims a downbeat at the half note level and a downbeat at the quarter note level, both of these claims are added to a list of claimed facts. The second onset claims an upbeat at the quarter note level and a downbeat at the half note level. Since the downbeat at the half note level is already the list of claimed facts it will not be counted again.

To reiterate the first paragraph of this section more concretely, we define the evaluation function \textsc{score}($R, R'$) which counts the number of claims that are shared by $R$ and $R'$ and divides that by the total number of claims in $R$. The precision is defined as the number of claims in $R$ that appear in the claims of $R^*$ as well, divided by the total number of claims in $R$, so:
\begin{equation}
\label{eq:precision}
\mathrm{precision} = \textsc{score}(R, R^*).
\end{equation}
Similarly, the recall is defined as:
\begin{equation}
\label{eq:recall}
\mathrm{recall} = \textsc{score}(R^*, R).
\end{equation}

Consider the example in figure \ref{fig:precision_error}. The current evaluation assigns a precision of $\frac{6}{7}$ and a recall of $\frac{6}{6}$ to the parse in figure \ref{fig:precision_error:a}. This implies that a total of $6$ claims are made by the gold-standard. These are: the four down- and upbeats at the quarter note level and one downbeat and one upbeat at the half note level. The parse in figure \ref{fig:precision_error:b} makes the same claims but also claims an upbeat at the eighth note level, which is incorrect.

We have not addressed the second issue yet. That is, we are not sure if the top level of $R$ corresponds to the top level of $R^*$. Either of those levels may have been added as a result of a different phase in $R$ and $R^*$. The only solution seems to be to consider three scenarios, illustrated by figure \ref{fig:phase_levels}: (1) No extra levels have been added and the levels in $R$ are consistent with the levels in $R^*$. (2) Let $R^*$ be figure \ref{fig:phase_levels:a} and $R$ be figure \ref{fig:phase_levels:c}, then a phase error resulted in an extra level in $R$. (3) Let $R^*$ be figure \ref{fig:phase_levels:c} and $R$ be figure \ref{fig:phase_error:a}, then a phase error resulted in one level less in $R$. 


\begin{figure}
\subfloat[]{
\label{fig:phase_levels:a}
\parbox{0.3\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .$\bullet$ ] [ .$\bullet$ ] ] 
}
}
\subfloat[]{
\label{fig:phase_levels:b}
\parbox{0.3\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$\bullet$ ] ] [ .$*$ ] ]
}
}
\subfloat[]{
\label{fig:phase_levels:c}
\parbox{0.3\linewidth}{
\Tree
[ .{$\frac{1}{1}$} [ .{$\frac{1}{2}$} [ .$*$ ] [ .$\bullet$ ] ] [ .{$\frac{1}{2}$} [ .$\bullet$ ] [ .$*$ ] ] ]
}
}
\caption{Extra levels added by phase errors}
\label{fig:phase_levels}
\end{figure}

If we want to get a valid precision and recall in scenario (2), we should convert $R$ to the tree in figure \ref{fig:phase_levels:b} by adding a top-level tie. In scenario (3) we want to convert $R^*$ to the tree in figure \ref{fig:phase_levels:b} by adding a top-level tie. This suggests three scenarios for evaluation: (1) evaluate $R$ against $R^*$, (2) evaluate $(R, *)$ against $R^*$ and (3) evaluate $R$ against $(R^*, *)$. 

The evaluation function in its final form is:
\begin{align}
\label{eq:evaluation}
&\mathrm{precision} = \textsc{max}(\textsc{score}(R, R^*), \textsc{score}((R, *), R^*), \textsc{score}(R, (R^*, *))),\\
&\mathrm{recall} = \textsc{max}(\textsc{score}(R^*, R), \textsc{score}((R^*, *), R), \textsc{score}(R^*, (R, *))).
\end{align}


% Training and testsets

% Corpus and training and test set sizes (first 4 bars)


